Ion pairing controls rheological properties of polyelectrolyte crocodile line hydrogels

### 2016-01-14 PALSE/ion_pairing_paper.tex
apolar micro environement

Between two cross-links the crocodile chain can be considered as an entropic spring made of $N=n N_0$ monomers. The shear modulus of the gel is then given by:
\begin{equation}
G = \frac{c}{N}k_\mathrm{B}T,
\label{eq:G}
\end{equation}
where $c$ is the monomer number density, $k_\mathrm{B}$ is the Boltzmann constant and $T$ the temperature. Knowing the molecular mass $M$ of the chains, density $d$ of the solvent and the weight fraction $w$ of polymer, we obtain
\begin{equation}
n = \frac{\mathcal{N_A}}{M} w d \frac{k_\mathrm{B}T}{G}.
\end{equation}

@@

For \ce{PIm+Br-} we measure $\gamma_c = 800\%$ and thus $E_c \approx 65 k_\mathrm{B}T$ per head-to-body bond or \SI{160}{\kilo\joule\per\mol}. This value is much higher than the $E_i\approx\SI{5}{\kilo\joule\per\mol}$ typical of ionic bonds in water\cite{Schneider1992}, even when adding up the two bonds of the bivalent phosphonate head. This implies that the electrostatic bond is reinforced by other supramolecular interactions, probably hydrophobic interactions between the surrounding uncharged monomers of both chains and $\pi$-stacking of the imidazolium\cite{Rehm2010}.

@@

The spatial conformation of a polyelectrolyte chain in semi-dilute, low salt solution is a persistent random walk where the persistence length is the screening length $r_\mathrm{src}$ due to the dissociated counter ions~\cite{Rubinstein1996}. In bad solvent, the persistent random walk is made of electrostatic blobs of diameter $D$ containing $g_e$ monomers, see Figure~\ref{fig:dissociation}a:
\begin{align}
g_e &= \frac{A^2}{u}\tau\label{eq:ge},\\
D &= b \left(\frac{A^2}{u}\right)^{1/3}\label{eq:D}.
\end{align}
Here $A$ is the number of monomers between uncondensed charges, $\tau = 1 - T/\Theta$ the reduced temperature and $u = \ell_\mathrm{B}/b \approx 2.7$ is the ratio of Bjerrum length $\ell_\mathrm{B}$ and monomer size $b\approx \SI{0.26}{\nano\metre}$.

>>>>

Between two cross-links the crocodile chain can be considered as an entropic spring made of $N=n N_0$ monomers. The shear modulus of the gel is then given by\cite{Rubinstein1996}:
\begin{equation}
G = \frac{c}{N}k_\mathrm{B}T,
\label{eq:G}
\end{equation}
where $c$ is the monomer number density, $k_\mathrm{B}$ is the Boltzmann constant and $T$ the temperature. Knowing the molecular mass $M$ of the chains, density $d$ of the solvent and the weight fraction $w$ of polymer, we obtain
\begin{equation}
n = \frac{\mathcal{N_A}}{M} w d \frac{k_\mathrm{B}T}{G}.
\end{equation}

@@

For \ce{PIm+Br-} we measure $\gamma_c = 800\%$ and thus $E_c \approx 65 k_\mathrm{B}T$ per head-to-body bond or \SI{160}{\kilo\joule\per\mol}. This value is much higher than the $E_i\approx\SI{5}{\kilo\joule\per\mol}$ typical of ionic bonds in water\cite{Schneider1992}, even when adding up the two bonds of the bivalent phosphonate head. However the surrounding medium is not water but the hydrophobic chains. Such an apolar microenvironement is well known to enhance otherwise weak electrostatic interactions in protein folding or engineered self-assembly\cite{Rehm2010}. The local dielectric constant can easily be 16 times lower than water, justifying the large bonding energy we measure.

@@

The spatial conformation of a polyelectrolyte chain in semi-dilute, low salt solution is a persistent random walk where the persistence length is the screening length $r_\mathrm{src}$ due to the dissociated counter ions\cite{Rubinstein1996}. In bad solvent, the persistent random walk is made of electrostatic blobs of diameter $D$ containing $g_e$ monomers, see Figure~\ref{fig:dissociation}a:
\begin{align}
g_e &= \frac{A^2}{u}\tau\label{eq:ge},\\
D &= b \left(\frac{A^2}{u}\right)^{1/3}\label{eq:D}.
\end{align}
Here $A$ is the number of monomers between uncondensed charges, $\tau = 1 - T/\Theta$ the reduced temperature and $u = \ell_\mathrm{B}/b \approx 2.7$ is the ratio of Bjerrum length $\ell_\mathrm{B}$ and monomer size $b\approx \SI{0.26}{\nano\metre}$.
